---
title: 'R Graphics Cookbook: Chapter 6'
author: "Brandon Foltz"
date: "2023-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
```
```{r}
ggplot(faithful, aes(x = waiting)) +
  geom_histogram()

ggplot(faithful, aes(x = waiting)) +
  geom_histogram(binwidth = 5, fill = "white", color = "black")

binsize <- diff(range(faithful$waiting))/15

ggplot(faithful, aes(x = waiting)) +
  geom_histogram(binwidth = binsize, fill = "white", color = "black")
```

Multiple Histograms
```{r}
ggplot(birthwt, aes(x = bwt)) +
  geom_histogram(fill = "white", color = "black") +
  facet_grid(smoke ~ .)

birthwt_mod <- birthwt

birthwt_mod$smoke <- recode_factor(birthwt_mod$smoke, '0' = 'No Smoke', '1' = 'Smoke')
```

```{r}
ggplot(birthwt_mod, aes(x = bwt)) +
  geom_histogram(fill = "white", color = "black") +
  facet_grid(smoke ~ .)

ggplot(birthwt, aes(x = bwt)) +
  geom_histogram(fill = "white", color = "black") +
  facet_grid(race ~ .)

ggplot(birthwt, aes(x = bwt)) +
  geom_histogram(fill = "white", color = "black") +
  facet_grid(race ~., scales = "free")

ggplot(birthwt_mod, aes(x = bwt, fill = smoke)) +
  geom_histogram(position = "identity", alpha = 0.4)
```

Density Curve
```{r}
ggplot(faithful, aes(x = waiting)) +
  geom_density()

ggplot(faithful, aes(x = waiting)) +
  geom_line(stat = "density") +
  expand_limits(y = 0)
```
```{r}
ggplot(faithful, aes(x = waiting)) +
  geom_line(stat = "density") +
  geom_line(stat = "density", adjust = .25, color = "orange") + # less smooth
  geom_line(stat = "density", adjust = 2, color = "blue") # more smooth
```
```{r}
ggplot(faithful, aes(x = waiting)) +
  geom_density(fill = "blue", alpha = .2) +
  xlim(35, 105)

ggplot(faithful, aes(x = waiting)) +
  geom_density(fill = "blue", alpha = 0.2, color = NA) +
  xlim(35, 105) +
  geom_line(stat = "density")
```
To solve this problem, you can scale down the histogram to match the density curve with the mapping y = ..density... Here we’ll add geom_histogram() first, and then layer geom_density() on top.

```{r}
ggplot(faithful, aes(x = waiting, y = ..density..)) +
  geom_histogram(fill = "cornsilk", color = "grey60", size = 0.2) +
  geom_density() +
  xlim(35, 105)
```

Multiple Density Curves from Grouped Data

```{r}
birthwt_mod <- birthwt|>
  mutate(smoke = as.factor(smoke))

ggplot(birthwt_mod,
       aes(x = bwt, color = smoke)) +
  geom_density()

ggplot(birthwt_mod,
       aes(x = bwt, fill = smoke)) +
  geom_density(alpha = 0.3)
```
```{r}
ggplot(birthwt_mod,
       aes(x = bwt)) +
  geom_density() +
  facet_grid(smoke ~ .)

#levels(birthwt_mod$smoke)

birthwt_mod$smoke <- recode(birthwt_mod$smoke, '0' = 'No Smoke', '1' = 'Smoke')

ggplot(birthwt_mod, aes(x = bwt)) +
  geom_density() +
  geom_vline(xintercept = median(birthwt_mod$bwt)) +
  facet_grid(smoke ~ .)
```
```{r}
ggplot(birthwt_mod,
       aes(x = bwt, y = ..density..)) +
  geom_histogram(binwidth = 200, fill = "cornsilk", color = "grey60", size = 0.2) +
  geom_density() +
  facet_grid(smoke ~ .)
```

Frequency Polygon

```{r}
ggplot(faithful, aes(x = waiting)) +
  geom_freqpoly()

ggplot(faithful, aes(x = waiting)) +
  geom_freqpoly(binwidth = 4)

binsize <- diff(range(faithful$waiting))/15

ggplot(faithful, aes(x = waiting)) +
  geom_freqpoly(binwidth = binsize)
```

Box Plots
```{r}
ggplot(birthwt, aes(x = factor(race), y = bwt)) +
  geom_boxplot()

ggplot(birthwt, aes(x = factor(race), y = bwt)) +
  geom_boxplot(width = 0.5)

ggplot(birthwt, aes(x = factor(race), y = bwt)) +
  geom_boxplot(outlier.size = 1.5, outlier.shape = 21, outlier.color = "red")

ggplot(birthwt, aes(x = 1, y = bwt)) +
  geom_boxplot() +
  scale_x_continuous(breaks = NULL) +
  theme(axis.title.x = element_blank())
```

Notched Box Plot

```{r}
ggplot(birthwt, aes(x = factor(race), y = bwt)) +
  geom_boxplot(notch = TRUE)
```

Adding Means to a Boxplot

```{r}
ggplot(birthwt, aes(x = factor(race), y = bwt)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 3, fill = "red")
```

Violin Plot

```{r}
hw_p <- ggplot(heightweight, aes(x = sex, y = heightIn))

hw_p + geom_violin()

hw_p +
  geom_violin() +
  geom_boxplot(width = 0.1, fill = "black", outlier.colour = NA) +
  stat_summary(fun.y = median, geom = "point", fill = "white", shape = 21, size = 2.5)

hw_p + geom_violin(trim = FALSE)

hw_p + geom_violin(scale = "count")

hw_p + geom_violin(adjust = 2)

hw_p + geom_violin(adjust = .5)
```

Dot Plot

```{r}
c2009 <- countries|>
  filter(Year == 2009 & healthexp > 2000)

c2009_p <- ggplot(c2009, aes(infmortality))

c2009_p + geom_dotplot()

c2009_p + geom_dotplot(binwidth = 0.25) +
  geom_rug() +
  scale_y_continuous(breaks = NULL) +
  theme(axis.title.y = element_blank())

c2009_p +
  geom_dotplot(method = "histodot", binwidth = 0.25) +
  geom_rug() +
  scale_y_continuous(breaks = NULL) +
  theme(axis.title.y = element_blank())
```
Center stacking

```{r}
c2009_p +
  geom_dotplot(binwidth = 0.25, stackdir = "center") +
  scale_y_continuous(breaks = NULL) +
  theme(axis.title.y = element_blank())

c2009_p +
  geom_dotplot(binwidth = 0.25, stackdir = "centerwhole") +
  scale_y_continuous(breaks = NULL) +
  theme(axis.title.y = element_blank())
```

Multple Dot Plots

```{r}
ggplot(heightweight, aes(x = sex, y = heightIn)) +
  geom_dotplot(binaxis = "y", binwidth = 0.5, stackdir = "center")

ggplot(heightweight, aes(x = sex, y = heightIn)) +
  geom_boxplot(outlier.colour = NA, width = 0.4) +
  geom_dotplot(binaxis = "y", binwidth = 0.5, stackdir = "center", fill = NA)
```
```{r}
ggplot(heightweight, aes(x = sex, y = heightIn)) +
  geom_boxplot(aes(x = as.numeric(sex) + .2, group = sex), width = .25) +
  geom_dotplot(
    aes(x = as.numeric(sex) - .2, group = sex),
    binaxis = "y",
    binwidth = .5,
    stackdir = "center"
  ) +
  scale_x_continuous(
    breaks = 1:nlevels(heightweight$sex),
    labels = levels(heightweight$sex)
  )
```

Density Plot 2d

```{r}
faithful_p <- ggplot(faithful, aes(x = eruptions, y = waiting))

faithful_p +
  geom_point() +
  stat_density2d()

faithful_p +
  stat_density2d(aes(color = ..level..))

faithful_p +
  stat_density2d(aes(fill = ..density..), geom = "raster", contour = FALSE)

faithful_p + 
  geom_point() +
  stat_density2d(aes(fill = ..density..), geom = "raster", contour = FALSE)

faithful_p +
  geom_point() +
  stat_density2d(aes(alpha = ..density..), geom = "tile", contour = FALSE)

faithful_p +
  stat_density2d(
    aes(fill = ..density..),
    geom = "raster", 
    contour = FALSE,
    h = c(.5, 5)
  )
  
```

