---
title: 'R Graphics Cookbook: Chapter 8'
author: "Brandon Foltz"
date: "2023-03-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gcookbook)
library(MASS)
library(scales)
```
```{r}
# Flipping coordinates

ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot()

ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot() +
  coord_flip()

ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot() +
  coord_flip() +
  scale_x_discrete(limits = rev(levels(PlantGrowth$group)))
```

Setting the Range of a Continuous Axis

```{r}
pg_plot <- ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot()

pg_plot

pg_plot +
  ylim(0, max(PlantGrowth$weight))
```
Presedence

```{r}
pg_plot +
  ylim(0,10) +
  scale_y_continuous(breaks = NULL)

pg_plot +
  scale_y_continuous(breaks = NULL) +
  ylim(0,10)

pg_plot +
  scale_y_continuous(limits = c(0,10), breaks = NULL)
```
With the box plots in these examples, if you restrict the y range so that some of the original data is clipped, the box plot statistics will be computed based on clipped data, and the shape of the box plots will change.

With a coordinate transform, the data is not clipped; in essence, it zooms in or out to the specified range. Figure 8.4 shows the difference between the two methods:

```{r}
pg_plot +
  scale_y_continuous(limits = c(5, 6.5))

pg_plot +
  coord_cartesian(ylim = c(5, 6.5))

pg_plot +
  expand_limits(y = 0)
```

Reversing a Continuous Axis

```{r}
ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot() +
  scale_y_continuous()

ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot() +
  ylim(6.5, 3.5)

ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot() +
  scale_y_reverse(limits = c(8, 0))
```

Changing the Order of Items on a Categorical Axis

```{r}
pg_plot <- ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot()

pg_plot + scale_x_discrete(limits = c("trt1", "ctrl1", "trt2"))

# can also use this to disply only a subset of items
pg_plot + scale_x_discrete(limits = c("ctrl", "trt1"))

# reverse
pg_plot + scale_x_discrete(limits = rev(levels(PlantGrowth$group)))
```

Setting the Scaling Ratio of the X- and Y-Axes

```{r}
m_plot <- ggplot(marathon, aes(x = Half, y = Full)) +
  geom_point()

m_plot + coord_fixed()

m_plot +
  coord_fixed() +
  scale_y_continuous(breaks = seq(0, 420, 30)) +
  scale_y_continuous(breaks = seq(0, 420, 30))

m_plot +
  coord_fixed(ratio = 1/2) +
  scale_y_continuous(breaks = seq(0, 420, 30)) +
  scale_x_continuous(breaks = seq(0, 420, 15))
```
Setting the Positions of Tick Marks

```{r}
ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot()

ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot() +
  scale_y_continuous(breaks = c(4, 4.25, 4.5, 5, 6, 8))

ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot() +
  scale_x_discrete(limits = c("trt2", "ctrl"), breaks = "ctrl")
```

Removing Tick Marks and Labels

```{r}
pg_plot <- ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot()

pg_plot +
  theme(axis.text.x = element_blank())

pg_plot +
  theme(axis.ticks = element_blank(), axis.text.y = element_blank())

pg_plot +
  scale_y_continuous(breaks = NULL)
```

Changing the Text of Tick Labels

```{r}
hw_plot <- ggplot(heightweight, aes(x = ageYear, y = heightIn)) +
  geom_point()

hw_plot
```
```{r}
hw_plot +
  scale_y_continuous(
    breaks = c(50, 56, 60, 66, 72),
    labels = c("Tiny", "Really\nshort", "Short", "Medium", "Tallish")
  )
```
```{r}
footinch_formatter <- function(x){
  foot <- floor(x/12)
  inch <- x %% 12
  return(paste(foot, "'", inch, "\"", sep = ""))
}

footinch_formatter(56:64)

hw_plot +
  scale_y_continuous(breaks = seq(48, 72, 4), labels = footinch_formatter)
```
```{r}
timeHMS_formatter <- function(x){
  h <- floor(x/60)
  m <- floor(x %% 60)
  s <- round(60*(x %% 1))
  lab <- sprintf("%02d:%02d:%02d", h, m, s)
  lab <- gsub("^00:", "", lab)
  lab <- gsub("^0", "", lab)
  return(lab)
}
```
```{r}
timeHMS_formatter(c(.33, 50, 51.25, 59.32, 60, 60.1, 130.23))
```
The scales package, which is installed with ggplot2, comes with some built-in formatting functions:

comma() adds commas to numbers, in the thousand, million, billion, etc. places.
dollar() adds a dollar sign and rounds to the nearest cent.
percent() multiplies by 100, rounds to the nearest integer, and adds a percent sign.
scientific() gives numbers in scientific notation, like 3.30e+05, for large and small numbers.
If you want to use these functions, you must first load the scales package, with library(scales)

Changing the Appearance of Tick Labels

```{r}
pg_plot <- ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot() +
  scale_x_discrete(
    breaks = c("ctrl", "trt1", "trt2"),
    labels = c("Control", "Treatment 1", "Treatment 2")
  )

pg_plot

pg_plot +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

pg_plot +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1))
```

```{r}
pg_plot +
  theme(
    axis.text.x = element_text(family = "Times", face = "italic",
                               color = "darkred", size = rel(0.9))
  )
```

Changing the Text of Axis Labels

```{r}
hw_plot <- ggplot(heightweight, aes(x = ageYear, y = heightIn, color = sex)) +
  geom_point()

hw_plot

hw_plot +
  xlab("Age in years") +
  ylab("Height in inches")

hw_plot +
  labs(x = "Age in years", y = "Height in inches")

hw_plot +
  scale_x_continuous(name = "Age in years")

hw_plot +
  scale_x_continuous(name = "Age\n(years)")
```

Removing Axis Labels

```{r}
pg_plot <- ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot()

pg_plot +
  xlab(NULL)
```

Changing the Appearance of Axis Labels

```{r}
hw_plot <- ggplot(heightweight, aes(x = ageYear, y = heightIn)) +
  geom_point()

hw_plot +
  theme(axis.title.x = element_text(face = "italic", color = "darkred", size = 14))

hw_plot +
  ylab("Height\n(inches)") +
  theme(axis.title.y = element_text(angle = 0, face = "italic", size = 14))

hw_plot +
  ylab("Height\n(inches)") +
  theme(axis.title.y = element_text(
    angle = 90,
    face = "italic",
    color = "darkred",
    size = 14
  ))
```

Showing Lines Along the Axes

```{r}
hw_plot <- ggplot(heightweight, aes(x = ageYear, y = heightIn)) +
  geom_point()

hw_plot +
  theme(axis.line = element_line(color = "black"))

hw_plot +
  theme_bw() +
  theme(panel.border = element_blank(), axis.line = element_line(color = "black"))

hw_plot +
  theme_bw() +
  theme(
    panel.border = element_blank(),
    axis.line = element_line(color = "black", size = 4)
    )

hw_plot +
  theme_bw() +
  theme(
    panel.border = element_blank(),
    axis.line = element_line(color = "black", size = 4, lineend = "square")
  )
```

Using a Logarithmic Axis

```{r}
animals_plot <- ggplot(Animals, aes(x = body, y = brain, label = rownames(Animals))) +
  geom_text(size = 3)

animals_plot

animals_plot +
  scale_x_log10() +
  scale_y_log10()

animals_plot +
  scale_x_log10(breaks = 10^(-1:5)) +
  scale_y_log10(breaks = 10^(0:3))

ggplot(Animals, aes(x = log10(body), y = log10(brain), label = rownames(Animals))) +
  geom_text(size = 3)

# Use natural log on x, and log2 on y
animals_plot +
  scale_x_continuous(
    trans = log_trans(),
    breaks = trans_breaks("log", function(x) exp(x)),
    labels = trans_format("log", math_format(e^.x))
  ) +
  scale_y_continuous(
    trans = log2_trans(),
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x))
  )
```

Log Time Series

```{r}
ggplot(aapl, aes(x = date, y = adj_price)) +
  geom_line()

ggplot(aapl, aes(x = date, y = adj_price)) +
  geom_line() +
  scale_y_log10(breaks = c(2,10,50,250))
```

Adding Ticks for a Logarithmic Axis

```{r}
breaks_log10 <- function(x){
  low <- floor(log10(min(x)))
  high <- ceiling(log10(max(x)))
  
  10^(seq.int(low, high))
}

ggplot(Animals, aes(x = body, y = brain, label = rownames(Animals))) +
  geom_text(size = 3) +
  annotation_logticks() +
  scale_x_log10(breaks = breaks_log10,
                labels = trans_format(log10, math_format(10^.x))) +
  scale_y_log10(breaks = breaks_log10,
                labels = trans_format(log10, math_format(10^.x)))
```
```{r}
breaks_5log10 <- function(x){
  low <- floor(log10(min(x)/5))
  high <- ceiling(log10(max(x)/5))
  
  5 * 10^(seq.int(low, high))
}

ggplot(Animals, aes(x = body, y = brain, label = rownames(Animals))) +
  geom_text(size = 3) +
  annotation_logticks() +
  scale_x_log10(
    breaks = breaks_log10,
    minor_breaks = breaks_5log10,
    labels = trans_format(log10, math_format(10^.x))) +
  scale_y_log10(
    breaks = breaks_log10,
    minor_breaks = breaks_5log10,
    labels = trans_format(log10, math_format(10^.x))) +
  coord_fixed() +
  theme_bw()
```

Making a Circular Plot

```{r}
ggplot(wind, aes(x = DirCat, fill = SpeedCat)) +
  geom_histogram(binwidth = 15, boundary = -7.5) +
  coord_polar() +
  scale_x_continuous(limits = c(0, 360))

ggplot(wind, aes(x = DirCat, fill = SpeedCat)) +
  geom_histogram(binwidth = 15, boundary = -7.5, color = "black", size = 0.25) +
  guides(fill = guide_legend(reverse = TRUE)) +
  coord_polar() +
  scale_x_continuous(
    limits = c(0, 360),
    breaks = seq(0,360, by = 45),
    minor_breaks = seq(0, 360, by = 15)) +
  scale_fill_brewer()
```
```{r}
mdeaths_mod <- data.frame(
  deaths = as.numeric(mdeaths),
  month = as.numeric(cycle(mdeaths))
)

mdeaths_mod <- mdeaths_mod|>
  group_by(month)|>
  summarise(deaths = mean(deaths))

# mdeaths_mod

mdeaths_plot <- ggplot(mdeaths_mod,
                       aes(x = month, y = deaths)) +
  geom_line() +
  scale_x_continuous(breaks = 1:12)

mdeaths_plot +coord_polar()

mdeaths_plot +coord_polar() + ylim(0, NA)

mdeaths_plot +
  coord_polar() +
  ylim(0, max(mdeaths_mod$deaths)) +
  xlim(0, 12)

# Connect the lines by adding a value for 0 that is the same as 12
mdeaths_x <- mdeaths_mod[mdeaths_mod$month==12, ]
mdeaths_x$month <- 0
mdeaths_new <- rbind(mdeaths_x, mdeaths_mod)

# Make the same plot as before, but with the new data, by using %+%
mdeaths_plot %+%
  mdeaths_new +
  coord_polar() +
  ylim(0, NA)
```

Using Dates on an Axis

```{r}
ggplot(economics, aes(x = date, y = psavert)) +
  geom_line()

econ_mod <- economics|>
  filter(date >= as.Date("1992-05-01") & date < as.Date("1993-06-01"))

econ_plot <- ggplot(econ_mod, aes(x = date, y = psavert)) +
  geom_line()

econ_plot

datebreaks <- seq(as.Date("1992-06-01"), as.Date("1993-06-01"), by = "2 month")

econ_plot +
  scale_x_date(breaks = datebreaks) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

econ_plot +
  scale_x_date(breaks = datebreaks, labels = date_format("%Y %b")) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Using Relative Times on an Axis

```{r}
# Convert WWWusage time-series object to data frame
www <- data.frame(
  minute = as.numeric(time(WWWusage)),
  users  = as.numeric(WWWusage)
)

# Define a formatter function - converts time in minutes to a string
timeHM_formatter <- function(x) {
  h <- floor(x/60)
  m <- floor(x %% 60)
  lab <- sprintf("%d:%02d", h, m) # Format the strings as HH:MM
  return(lab)
}

# Default x axis
ggplot(www, aes(x = minute, y = users)) +
  geom_line()

# With formatted times
ggplot(www, aes(x = minute, y = users)) +
  geom_line() +
  scale_x_continuous(
    name = "time",
    breaks = seq(0, 100, by = 10),
    labels = timeHM_formatter
  )
```
```{r}
scale_x_continuous(
  breaks = c(0, 20, 40, 60, 80, 100),
  labels = c("0:00", "0:20", "0:40", "1:00", "1:20", "1:40")
)

timeHMS_formatter <- function(x) {
  h <- floor(x/3600)
  m <- floor((x/60) %% 60)
  s <- round(x %% 60)                       # Round to nearest second
  lab <- sprintf("%02d:%02d:%02d", h, m, s) # Format the strings as HH:MM:SS
  lab <- sub("^00:", "", lab)               # Remove leading 00: if present
  lab <- sub("^0", "", lab)                 # Remove leading 0 if present
  return(lab)
}

timeHMS_formatter(c(20, 3000, 3075, 3559.2, 3600, 3606, 7813.8))
```

