---
title: 'R Graphics Cookbook: Chapter 12'
author: "Brandon Foltz"
date: "2023-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
library(gcookbook)
library(viridis)
library(RColorBrewer)
library(scales)
```
Setting the colors of objects
```{r}
ggplot(birthwt, aes(x = bwt)) +
  geom_histogram(fill = "orange", color = "black")

ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(color = "orange")
```

Representing Variables with Color

```{r}
ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar)) +
  geom_col(color = "black", position = "dodge")

ggplot(cabbage_exp, aes(x = Date, y = Weight)) +
  geom_col(aes(fill = Cultivar), color = "black", position = "dodge")

ggplot(mtcars, aes(x = wt, y = mpg, color = cyl)) +
  geom_point()

ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(aes(color = cyl))
```
Converting to a Factor
```{r}
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point()

mtcars_mod <- mtcars|>
  mutate(cyl = as.factor(cyl))

ggplot(mtcars_mod, aes(x = wt, y = mpg, color = cyl)) +
  geom_point()
```

Using a Colorlind-Friendly Palette (Viridis)

You will need to add scale_fill_viridis_c() to your plot if your data is continuous. If your data is discrete you will need to use scale_fill_viridis_d() instead.

```{r}
uspopage_plot <- ggplot(uspopage, aes(x = Year, y = Thousands, fill = AgeGroup)) +
  geom_area()

uspopage_plot + scale_fill_viridis_d()
```

Using a Different Palette for a Discrete Variable

```{r}
uspopage_plot <- ggplot(uspopage, aes(x = Year, y = Thousands, fill = AgeGroup)) +
  geom_area()

uspopage_plot

uspopage_plot +
  scale_fill_viridis(discrete = TRUE)

uspopage_plot +
  scale_fill_brewer()
```

```{r}
hw_splot <- ggplot(heightweight, aes(x = ageYear, y = heightIn, color = sex)) +
  geom_point()

hw_splot

hw_splot +
  scale_color_hue(l = 45)
```
```{r}
display.brewer.all()
```
```{r}
hw_splot +
  scale_color_brewer(palette = "Oranges") +
  theme_bw()

hw_splot +
  scale_color_grey()

hw_splot +
  scale_color_grey(start = 0.7, end = 0)
```

Using a Manually Defined Palette for a Discrete Variable

```{r}
hw_plot <- ggplot(heightweight, aes(x = ageYear, y = heightIn, color = sex)) +
  geom_point()

hw_plot + scale_color_manual(values = c("orange", "lightblue"))

hw_plot + scale_color_manual(values = c("#CC6666", "#7777DD"))

hw_plot + scale_color_manual(values = c("#440154FF", "#FDE725FF")) +
  theme_bw()

hw_plot + scale_color_manual(values = c(m = "lightblue", f = "orange"))
```
```{r}
viridis(2)
inferno(5)
```
Using a Manually Defined Palette for a Continuous Variable

```{r}
hw_plot <- ggplot(heightweight, aes(x = ageYear, heightIn, color = weightLb)) +
  geom_point(size = 3)

hw_plot

hw_plot +
  scale_color_gradient2(
    low = muted("red"),
    mid = "white",
    high = muted("blue"),
    midpoint = 110
  )

hw_plot +
  scale_color_gradient(low = "black", high = "white")

hw_plot +
  scale_color_gradientn(colors = c("darkred", "orange", "yellow", "white"))
```
Coloring a Shaded Region Based on Value

```{r}
climate_mod <- climate|>
  filter(Source == "Berkeley")|>
  mutate(valence = if_else(Anomaly10y >= 0, "pos", "neg"))

climate_mod

ggplot(climate_mod, aes(x = Year, y = Anomaly10y)) +
  geom_area(aes(fill = valence)) +
  geom_line() +
  geom_hline(yintercept = 0)
```
```{r}
interp <- approx(climate_mod$Year, climate_mod$Anomaly10y, n = 1000)

cbi <- data.frame(Year = interp$x, Anomaly10y = interp$y)|>
  mutate(valence = if_else(Anomaly10y >= 0, "pos", "neg"))

ggplot(cbi, aes(x = Year, y = Anomaly10y)) +
  geom_area(aes(fill = valence), alpha = 0.4) +
  geom_line() +
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = c("#CCEEFF", "#FFDDDD"), guide = "none") +
  scale_x_continuous(expand = c(0,0)) +
  theme_bw()
```


