---
title: 'R Graphics Cookbook: Misc Graphs'
author: "Brandon Foltz"
date: "2023-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(corrplot)
library(igraph)
library(gcookbook)
library(rgl)
library(maps)
```

```{r}
mcor <- cor(mtcars)
round(mcor, digits = 2)

corrplot(mcor)
corrplot(mcor, method = "shade", shade.col = NA, tl.col = "black", tl.srt = 45)

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(mcor, method = "shade", shade.col = NA, tl.col = "black", tl.srt = 45,
         col = col(200), addCoef.col = "black", cl.pos = "n", order = "AOE")
```
Plotting a Function

```{r}
p <- ggplot(data.frame(x = c(-3,3)), aes(x=x))
p + stat_function(fun = dnorm)
p + stat_function(fun = dt, args = list(df = 2))
```
```{r}
myfun <- function(xvar){
  1 / (1 + exp(-xvar + 10))
}

ggplot(data.frame(x = c(0, 20)), aes(x=x)) +
  stat_function(fun = myfun)
```
Shading a Subregion Under a Function Curve

```{r}
dnorm_limit <- function(x){
  y <- dnorm(x)
  y[x < 0 | x > 2]  <- NA
  return(y)
}

p <- ggplot(data.frame(x = c(-3,3)), aes(x=x))

p +
  stat_function(fun = dnorm_limit, geom = "area", fill = "blue", alpha = 0.2) +
  stat_function(fun = dnorm)
```
```{r}
limitRange <- function(fun, min, max){
  function(x){
    y <- fun(x)
    y[x < min | x > max] <- NA
    return(y)
  }
}

dlimit <- limitRange(dnorm, 0, 2)
dlimit(-2:4)

p + 
  stat_function(fun = dnorm) +
  stat_function(fun = limitRange(dnorm, 0, 2), geom = "area", fill = "blue", alpha = 0.2)
```

Creating a Network Graph

```{r}
gd <- graph(c(1,2, 2,3, 2,4, 1,4, 5,5, 3,6))
plot(gd)

gu <- graph(c(1,2, 2,3, 2,4, 1,4, 5,5, 3,6), directed = FALSE)
plot(gu, vertex.label = NA)

set.seed(229)
plot(gu)
```

```{r}
madmen2

g <- graph.data.frame(madmen2, directed = TRUE)

par(mar = c(0,0,0,0))
plot(g, layout = layout.fruchterman.reingold, vertex.size = 8, edge.arrow.size = 0.5, vertex.label = NA)
```


```{r}
g <- graph.data.frame(madmen, directed = FALSE)
par
```

Using Text Labels in a Network Graph

```{r}
# Copy madmen and drop every other row
m <- madmen[1:nrow(madmen) %% 2 == 1, ]

g <- graph.data.frame(m, directed = FALSE)

V(g)$name

plot(g, layout = layout.fruchterman.reingold,
     vertex.size = 4,
     vertex.label = V(g)$name,
     vertex.label.cex = 0.8,
     vertex.label.dist = 0.4,
     vertex.label.color = "black")

# View the edges
E(g)
```

Creating a Heat Map

```{r}
presidents

# convert to data frame for ggplot

pres_rating <- data.frame(
  rating = as.numeric(presidents),
  year = as.numeric(floor(time(presidents))),
  quarter = as.numeric(cycle(presidents))
)

pres_rating
```
```{r}
p  <- ggplot(pres_rating, aes(x = year, y = quarter, fill = rating))

p + geom_tile()

p + geom_raster()
```
```{r}
p +
  geom_tile() +
  scale_x_continuous(breaks = seq(1940, 1976, by = 4), expand = c(0,0)) +
  scale_y_reverse(expand = c(0,0)) +
  scale_fill_gradient2(midpoint = 50, mid = "grey70", limits = c(0, 100))
```
 Creating a 3D Scatter Plot
 
```{r}
plot3d(mtcars$wt, mtcars$disp, mtcars$mpg, type = "s", size = 0.75, lit = FALSE)
```
 
```{r}
# Function to interleave the elements of two vectors
interleave <- function(v1, v2)  as.vector(rbind(v1,v2))

# Plot the points
plot3d(mtcars$wt, mtcars$disp, mtcars$mpg,
       xlab = "Weight", ylab = "Displacement", zlab = "MPG",
       size = .75, type = "s", lit = FALSE)

# Add the segments
segments3d(interleave(mtcars$wt,   mtcars$wt),
           interleave(mtcars$disp, mtcars$disp),
           interleave(mtcars$mpg,  min(mtcars$mpg)),
           alpha = 0.4, col = "blue")
```
```{r}
# Make plot without axis ticks or labels
plot3d(mtcars$wt, mtcars$disp, mtcars$mpg,
       xlab = "", ylab = "", zlab = "",
       axes = FALSE,
       size = .75, type = "s", lit = FALSE)

segments3d(interleave(mtcars$wt,   mtcars$wt),
           interleave(mtcars$disp, mtcars$disp),
           interleave(mtcars$mpg,  min(mtcars$mpg)),
           alpha = 0.4, col = "blue")

# Draw the box.
rgl.bbox(color = "grey50",          # grey60 surface and black text
         emission = "grey50",       # emission color is grey50
         xlen = 0, ylen = 0, zlen = 0)  # Don't add tick marks

# Set default color of future objects to black
rgl.material(color = "black")

# Add axes to specific sides. Possible values are "x--", "x-+", "x+-", and "x++".
axes3d(edges = c("x--", "y+-", "z--"),
       ntick = 6,                       # Attempt 6 tick marks on each side
       cex = .75)                       # Smaller font

# Add axis labels. 'line' specifies how far to set the label from the axis.
mtext3d("Weight",       edge = "x--", line = 2)
mtext3d("Displacement", edge = "y+-", line = 3)
mtext3d("MPG",          edge = "z--", line = 3)
```

```{r}
# Given a model, predict zvar from xvar and yvar
# Defaults to range of x and y variables, and a 16x16 grid
predictgrid <- function(model, xvar, yvar, zvar, res = 16, type = NULL) {
  # Find the range of the predictor variable. This works for lm and glm
  # and some others, but may require customization for others.
  xrange <- range(model$model[[xvar]])
  yrange <- range(model$model[[yvar]])

  newdata <- expand.grid(x = seq(xrange[1], xrange[2], length.out = res),
                         y = seq(yrange[1], yrange[2], length.out = res))
  names(newdata) <- c(xvar, yvar)
  newdata[[zvar]] <- predict(model, newdata = newdata, type = type)
  newdata
}


# Convert long-style data frame with x, y, and z vars into a list
# with x and y as row/column values, and z as a matrix.
df2mat <- function(p, xvar = NULL, yvar = NULL, zvar = NULL) {
  if (is.null(xvar)) xvar <- names(p)[1]
  if (is.null(yvar)) yvar <- names(p)[2]
  if (is.null(zvar)) zvar <- names(p)[3]

  x <- unique(p[[xvar]])
  y <- unique(p[[yvar]])
  z <- matrix(p[[zvar]], nrow = length(y), ncol = length(x))

  m <- list(x, y, z)
  names(m) <- c(xvar, yvar, zvar)
  m
}

# Function to interleave the elements of two vectors
interleave <- function(v1, v2)  as.vector(rbind(v1,v2))

library(rgl)

# Make a copy of the data set
m <- mtcars

# Generate a linear model
mod <- lm(mpg ~ wt + disp + wt:disp, data = m)

# Get predicted values of mpg from wt and disp
m$pred_mpg <- predict(mod)

# Get predicted mpg from a grid of wt and disp
mpgrid_df <- predictgrid(mod, "wt", "disp", "mpg")
mpgrid_list <- df2mat(mpgrid_df)

# Make the plot with the data points
plot3d(m$wt, m$disp, m$mpg, type = "s", size = 0.5, lit = FALSE)

# Add the corresponding predicted points (smaller)
spheres3d(m$wt, m$disp, m$pred_mpg, alpha = 0.4, type = "s", size = 0.5, lit = FALSE)

# Add line segments showing the error
segments3d(interleave(m$wt,   m$wt),
           interleave(m$disp, m$disp),
           interleave(m$mpg,  m$pred_mpg),
           alpha = 0.4, col = "red")

# Add the mesh of predicted values
surface3d(mpgrid_list$wt, mpgrid_list$disp, mpgrid_list$mpg,
          alpha = 0.4, front = "lines", back = "lines")
```
```{r}
plot3d(mtcars$wt, mtcars$disp, mtcars$mpg,
       xlab = "", ylab = "", zlab = "",
       axes = FALSE,
       size = .5, type = "s", lit = FALSE)

# Add the corresponding predicted points (smaller)
spheres3d(m$wt, m$disp, m$pred_mpg, alpha = 0.4, type = "s", size = 0.5, lit = FALSE)

# Add line segments showing the error
segments3d(interleave(m$wt,   m$wt),
           interleave(m$disp, m$disp),
           interleave(m$mpg,  m$pred_mpg),
           alpha = 0.4, col = "red")

# Add the mesh of predicted values
surface3d(mpgrid_list$wt, mpgrid_list$disp, mpgrid_list$mpg,
          alpha = 0.4, front = "lines", back = "lines")

# Draw the box
rgl.bbox(color = "grey50",          # grey60 surface and black text
         emission = "grey50",       # emission color is grey50
         xlen = 0, ylen = 0, zlen = 0)  # Don't add tick marks

# Set default color of future objects to black
rgl.material(color = "black")

# Add axes to specific sides. Possible values are "x--", "x-+", "x+-", and "x++".
axes3d(edges = c("x--", "y+-", "z--"),
       ntick = 6,                       # Attempt 6 tick marks on each side
       cex = .75)                       # Smaller font

# Add axis labels. 'line' specifies how far to set the label from the axis.
mtext3d("Weight",       edge = "x--", line = 2)
mtext3d("Displacement", edge = "y+-", line = 3)
mtext3d("MPG",          edge = "z--", line = 3)
```

Animating a 3D plot

```{r}
library(rgl)
plot3d(mtcars$wt, mtcars$disp, mtcars$mpg, type = "s", size = 0.75, lit = FALSE)

play3d(spin3d())
```
```{r}
# Spin on x-axis, at 4 rpm, for 20 seconds
play3d(spin3d(axis = c(1,0,0), rpm = 4), duration = 20)
```
```{r}
# Spin on z axis, at 4 rpm, for 15 seconds
movie3d(spin3d(axis = c(0,0,1), rpm = 4), duration = 15, fps = 50)
```

Creating a Dendrogram

```{r}
set.seed(392)

c2 <- countries |>
  filter(Year == 2009)|>
  drop_na()|>
  sample_n(25)

c2

rownames(c2) <- c2$Name
c2 <- c2[, 4:7]

c3 <- scale(c2)
c3
```
```{r}
hc <- hclust(dist(c3))
plot(hc)
plot(hc, hang = -1)
```

Creating a Vector Field

```{r}
isabel

islice <- filter(isabel, z == min(z))

ggplot(islice,
        aes(x = x, y = y)) +
  geom_segment(aes(xend = x + vx/50, yend = y + vy/50),
               linewidth = 0.25)
```
```{r}
islice <- filter(isabel, z == min(z))

every_n <- function(x, by = 2){
  x <- sort(x)
  x[seq(1, length(x), by = by)]
}

keepx <- every_n(unique(isabel$x), by = 4)
keepy <- every_n(unique(isabel$y), by = 4)

islicesub <- filter(islice, x %in% keepx & y %in% keepy)

library(grid)

ggplot(islicesub, aes(x=x, y=y)) +
  geom_segment(aes(xend = x+vx/50, yend = y+vy/50),
               arrow = arrow(length = unit(0.1, "cm")), size = 0.25)
```

```{r}
# The existing 'speed' column includes the z component. We'll calculate
# speedxy, the horizontal speed.
islicesub$speedxy <- sqrt(islicesub$vx^2 + islicesub$vy^2)

# Map speed to alpha
ggplot(islicesub, aes(x = x, y = y)) +
    geom_segment(aes(xend = x+vx/50, yend = y+vy/50, alpha = speed),
                 arrow = arrow(length = unit(0.1,"cm")), size = 0.6)
```
```{r}
# Get USA map data
usa <- map_data("usa")

# Map speed to colour, and set go from "grey80" to "darkred"
ggplot(islicesub, aes(x = x, y = y)) +
    geom_segment(aes(xend = x+vx/50, yend = y+vy/50, colour = speed),
                 arrow = arrow(length = unit(0.1,"cm")), size = 0.6) +
    scale_colour_continuous(low = "grey80", high = "darkred") +
    geom_path(aes(x = long, y = lat, group = group), data = usa) +
    coord_cartesian(xlim = range(islicesub$x), ylim = range(islicesub$y))
```
```{r}
# Keep 1 out of every 5 values in x and y, and 1 in 2 values in z
keepx <- every_n(unique(isabel$x), by = 5)
keepy <- every_n(unique(isabel$y), by = 5)
keepz <- every_n(unique(isabel$z), by = 2)

isub <- filter(isabel, x %in% keepx  &  y %in% keepy  &  z %in% keepz)

ggplot(isub, aes(x = x, y = y)) +
    geom_segment(aes(xend = x+vx/50, yend = y+vy/50, colour = speed),
                 arrow = arrow(length = unit(0.1,"cm")), size = 0.5) +
    scale_colour_continuous(low = "grey80", high = "darkred") +
    facet_wrap( ~ z)
```

Q-Q Plot

```{r}
ggplot(heightweight, aes(sample = heightIn)) +
  geom_qq() +
  geom_qq_line()

ggplot(heightweight, aes(sample = ageYear)) +
  geom_qq() +
  geom_qq_line()
```
Empirical Cumulative Distribution Function

```{r}
ggplot(heightweight, aes(x = heightIn)) +
  stat_ecdf()

ggplot(heightweight, aes(x = ageYear)) +
  stat_ecdf()
```
Creating a Mosaic Plot

```{r}
UCBAdmissions

library(vcd)

mosaic(~ Admit + Gender + Dept, data = UCBAdmissions)

mosaic(~ Dept + Gender + Admit, data = UCBAdmissions,
       highlighting = "Admit", highlighting_fill = c("lightblue", "pink"),
       direction = c("v", "h", "v"))
```

Creating a Map

```{r}
states_map <- map_data("state")

ggplot(states_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", color = "black")

ggplot(states_map, aes(x = long, y = lat, group = group)) +
  geom_path() + coord_map("mercator")
```
```{r}
world_map <- map_data("world")
world_map
sort(unique(world_map$region))
```
```{r}
east_asia <- map_data("world", region = c("Japan", "China", "North Korea", "South Korea"))

ggplot(east_asia, aes(x = long, y = lat, group = group, fill = region)) +
  geom_polygon(color = "black") +
  scale_fill_brewer(palette = "Set2")
```
```{r}
nz1 <- map_data("world", region = "New Zealand")|>
  filter(long > 0, lat > -48)

ggplot(nz1, aes(x = long, y = lat, group = group)) +
  geom_path()

nz2 <- map_data("nz")
ggplot(nz2, aes(x = long, y = lat, group = group)) +
  geom_path()
```

Choropleth Map

```{r}
crimes <- data.frame(state = tolower(rownames(USArrests)), USArrests)
crimes
```
```{r}
library(maps)
states_map <- map_data("state")
crime_map <- merge(states_map, crimes, by.x = "region", by.y = "state")
crime_map
crime_map <- arrange(crime_map, group, order)
crime_map

ggplot(crime_map, aes(x = long, y = lat, group = group, fill = Assault)) +
  geom_polygon(color = "black") +
  coord_map("polyconic")
```
```{r}
crime_p <- ggplot(crimes, aes(map_id = state, fill = Assault)) +
  geom_map(map = states_map, color = "black") +
  expand_limits(x = states_map$long, y = states_map$lat) +
  coord_map("polyconic")

crime_p +
  scale_fill_gradient2(low = "#559999", mid = "grey90", high = "#BB650B",
                       midpoint = median(crimes$Assault))

crime_p +
  scale_fill_viridis_c()
```
Use Quantiles for Discrete

```{r}
qa <- quantile(crimes$Assault, c(0, .2, .4, .6, .8, 1))
qa

crimes$Assault_q <- cut(crimes$Assault, qa,
                        labels = c("0 - 20%", "20 - 40%", "40 - 60%", "60 - 80%", "80 - 100%"),
                        include.lowest = TRUE)
crimes

pal <- colorRampPalette(c("#559999", "grey80", "#BB650B"))(5)
pal

ggplot(crimes, aes(map_id = state, fill = Assault_q)) +
  geom_map(map = states_map, color = "black") +
  scale_fill_manual(values = pal) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  coord_map("polyconic") +
  labs(fill = "Assault\nPercentile")
```
```{r}
ggplot(crimes, aes(map_id = state, fill = Assault)) +
  geom_map(map = states_map, color = "white") +
  expand_limits(x = states_map$long, y = states_map$lat) +
  coord_map("polyconic")
```
```{r}
ggplot(crimes, aes(map_id = state, fill = Assault_q)) +
  geom_map(map = states_map, color = "black") +
  scale_fill_manual(values = pal) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  coord_map("polyconic") +
  labs(fill = "Assualt\nPercentile") +
  theme_void()
```
```{r}
library(sf)
taiwan_shp <- st_read("fig/TWN_adm/TWN_adm2.shp")
ggplot(taiwan_shp) +
  geom_sf()
```

