---
title: 'R Graphics Cookbook: Chapter 15'
author: "Brandon Foltz"
date: "2023-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gcookbook)
```
```{r}
str(heightweight)
```
Creating a Dataframe
```{r}
g <- c("A", "B", "C")
x <- 1:3
dat <- data.frame(g, x)
dat

lst <- list(group = g, value = x)
dat <- as.data.frame(lst)

data_frame(g, x)

as_data_frame(lst)

as_tibble(dat)
```
Getting Information About Data Structure
```{r}
str(ToothGrowth)
summary(ToothGrowth)
```
```{r}
tg <- ToothGrowth
tg$supp <- as.character(tg$supp)
str(tg)
```
```{r}
ToothGrowth$supp
tg$supp
```
Adding a Column to a Dataframe
```{r}
ToothGrowth|>
  mutate(newcol = NA)

vec <- rep(c(1,2), 30)
ToothGrowth|>
  mutate(newcol = vec)
```
Deleting a Column from a Dataframe
```{r}
ToothGrowth|>
  select(-len, -supp)
```
Renaming Columns in a Dataframe
```{r}
ToothGrowth|>
  rename(length = len,
         supplement_type = supp)
```
Reordering Columns in a Dataframe
```{r}
ToothGrowth|>
  select(dose, len, supp)

ToothGrowth|>
  select(dose, everything())
```
Getting a Subset of a Dataframe
```{r}
library(gcookbook)
#climate

climate|>
  filter(Source == "Berkeley" & Year >= 1900 & Year <= 2000)

climate|>
  select(Year, Anomaly10y)

climate|>
  filter(Source == "Berkeley" & Year >= 1900 & Year <= 2000)|>
  select(Year, Anomaly10y)

slice(climate, 1:100)
```
Changing the Order of Factor Levels
```{r}
sizes <- factor(c("small", "large", "large", "small", "medium"))
sizes

factor(sizes, levels = c("small", "medium", "large"))

factor(c("small", "large", "large", "small", "medium"),
       levels = c("small", "medium", "large"))
```
Ordering a Factor
```{r}
factor(sizes, levels = rev(levels(sizes)))

fct_relevel(sizes, "small", "medium", "large")
```
Changing the Order of Factor Levels Based on Data Values
```{r}
iss <- InsectSprays
iss$spray

iss$spray <- reorder(iss$spray, iss$count, FUN = mean)
iss$spray

fct_reorder(iss$spray, iss$count, .fun = mean)
```
Changing the Name of Factor Levels
```{r}
sizes <- factor(c("small", "large", "large", "small", "medium"))
sizes

fct_recode(sizes, S = "small", M = "medium", L = "large")

do.call(
  fct_recode,
  c(list(sizes), setNames(c("small", "medium", "large"), c("S", "M", "L")))
)
```
```{r}
sizes <- factor(c("small", "large", "large", "small", "medium"))
levels(sizes) <- list(S = "small", M = "medium", L = "large")
sizes
```
Drop Unused Levels
```{r}
sizes <- factor(c("small", "large", "large", "small", "medium"))
sizes <- sizes[1:3]
sizes

droplevels(sizes)
fct_drop(sizes)
```

Changing the Name of Items in a Character Vector

```{r}
library(dplyr)

sizes <- c("small", "large", "large", "small", "medium")
sizes

# With recode(), pass it a named vector with the mappings
recode(sizes, small = "S", medium = "M", large = "L")

# Can also use quotes -- useful if there are spaces or other strange characters
recode(sizes, "small" = "S", "medium" = "M", "large" = "L")

do.call(
  recode,
  c(list(sizes), setNames(c("S", "M", "L"), c("small", "medium", "large")))
)
```
Recoding a Categorical Variable to Another Categorical Variable
```{r}
pg <- PlantGrowth[c(1,2,11,21,22), ]
pg

pg$treatment <- recode(pg$group, ctrl = "No", trt1 = "Yes", trt2 = "Yes")

fct_recode(pg$group, No = "ctrl", Yes = "trt1", Yes = "trt2")

pg$weightgroup <- interaction(pg$weight, pg$group)
pg
```
Recoding a Continuous Variable to a Categorical Variable

```{r}
pg <- PlantGrowth
pg$wtclass <- cut(pg$weight, breaks = c(0,5,6,Inf),
                  labels = c("small", "medium", "large"))
pg
```
Calculating New Columns From Existing Columns
```{r}
heightweight

heightweight|>
  mutate(heightCm = heightIn * 2.54,
         weightKg = weightLb * 2.204,
         bmi = weightKg / (heightCm / 100)^2)
```

Calculating New Columns by Groups

```{r}
library(MASS)

cabbages|>
  group_by(Cult)|>
  mutate(DevWt = HeadWt - mean(HeadWt))

mutate(cabbages, DevWt = HeadWt - mean(HeadWt))

cb <- cabbages|>
  group_by(Cult)|>
  mutate(DevWt = HeadWt - mean(HeadWt))
cb
```
```{r}
# The data before normalizing
ggplot(cb, aes(x = Cult, y = HeadWt)) +
  geom_boxplot()

# After normalizing
ggplot(cb, aes(x = Cult, y = DevWt)) +
  geom_boxplot()
```
You can also group the data frame on multiple variables and perform operations on multiple variables. The following code groups the data by Cult and Date, forming a group for each distinct combination of the two variables. After forming these groups, the code will calculate the deviation of HeadWt and VitC from the mean of each group:

```{r}
cabbages|>
  group_by(Cult, Date)|>
  mutate(
    DevWt = HeadWt - mean(HeadWt),
    DevVitC = VitC - mean(VitC)
  )
```

Summarizing Data by Groups

```{r}
cabbages|>
  group_by(Cult, Date)|>
  summarise(
    Weight = mean(HeadWt),
    VitC = mean(VitC)
  )
```
```{r}
cabbages|>
  group_by(Cult, Date)|>
  summarise(
    Weight = mean(HeadWt),
    sd = sd(HeadWt),
    n = n()
  )
```

Converting a Time Series Object to Times and Values

```{r}
nhtemp
as.numeric(time(nhtemp))
as.numeric(nhtemp)

nht <- data.frame(year = as.numeric(time(nhtemp)), temp = as.numeric(nhtemp))
nht
```
```{r}
pres_rating <- data.frame(
  year = as.numeric(time(presidents)),
  rating = as.numeric(presidents)
)
pres_rating

pres_rating2 <- data.frame(
  year = as.numeric(floor(time(presidents))),
  quarter = as.numeric(cycle(presidents)),
  rating = as.numeric(presidents)
)
pres_rating2
```

